---
title: "Danish Kings exercise"
author: "Kristian Krabbe"
date: "2023-10-06"
output: html_document
---
### 1. Opgave

**Calculate the average length of the reign for all monarchs**
a. You first need to know how to subset the relevant "Duration" column before you run a mean() on it.
b. Are you getting an error that there are characters in your column? Coerce your data to numbers with
as.numeric()
c. Remember to handle NAs: mean(X, na.rm=TRUE)
d. Want to do both tasks in one? Remember that you can cluster R functions: mean(as.numeric(X),
na.rm=TRUE)
```{r}
library(tidyverse)
read_csv("data/Dansk kongerække - Ark1.csv")

Kings <- read_csv("data/Dansk kongerække - Ark1.csv")
Kings <- Kings %>%
  mutate(Regeringstid = Regeringslut - Regeringsstart)
colnames(Kings)


mean(Kings$Regeringstid,na.rm = T)
```

### 2. Opgave
**How many kings enjoyed a longer than the average duration of reign?**
a. Select kings' names and Duration columns with select()
b. Filter() the kings dataset Duration > average_length and count the resulting rows with count()
c. Use mutate() to create a new column AboveAverage with TRUE/FALSE response depending on
whether the Duration is bigger or smaller than average_length
d. Group the unique values in this new column with group_by() and then tally(). What is your result?

```{r}
Kings %>% 
  select(Regent, Regeringstid) %>% 
  filter(Regeringstid > 19.3) %>% 
  count()  

Kings %>%  
mutate(OverGennemsnit = Regeringstid >19.3)
```
### 3. Opgave
**Sort kings by reign duration in descending order. Which three ruled the longest?**

```{r}
Kings %>% 
  arrange(desc(Regeringstid))
```
### 4. Opgave
**How many days did the three longest-ruling monarchs rule?**
a. Use mutate() to create a column called Days calculating the total number of days they ruled
b. BONUS: consider transition year (with 366 days)

```{r}
Kings %>% 
  mutate(Regeringsdage = Regeringstid * 365.25) %>% 
  arrange(desc(Regeringsdage))
```

### 5. Opgave
**How many Kristians/Christians or Frederiks were there among Danish kings?**
a. Search in the Names column via grepl() or str_detect()
b. Filter by this query and create a new table just with these kings.
  i. more complex query with combined filter(x == str_subset(x, pattern)) for very advanced
students

```{r}
Kings %>% 
  str_detect(Christian, Frederik)
```

